---
title: "Assignment 1"
author: "Max"
format: html
editor: visual
---

## Questions and Tasks 

1.  To begin, I opened a new version controlled project with Rstudio, choosing the Git option to clone the repository given in the assignment document. After that, I added packages to the project library and check / set working directory if needed.

```{r}
#library in package, packages are already installed
library(tidyverse)
library(dplyr)
library(haven)
getwd() #check working directory, it is set to assignment folder I created using point and click GIT/ project wizard, so don't need to change using setwd()


```

Now that the repository is available, I import the data and inspect for missing and unusual values. To inspect the data I first use the head function, and the view function to look at the entire data set.

```{r}
Avengers_data <- read.csv("avengers.csv")
head(Avengers_data)
view(Avengers_data)
```

2.  Now, to look at complete clean cases and create a subset with only these complete cases , will inspect for missing values and drop any rows with missing values.

    ```{r}
    clean_avengers <- drop_na(Avengers_data) #creating new object where rows with missing values are dropped using drop_na function in tidyverse
    view(clean_avengers) # new object looks as it should, missing values dropped. 

    ```

Next, the variable combat effectiveness is calculated as the sum of agility, speed, strength, and willpower

```{r}
avengers_effectiveness <- mutate(clean_avengers,combateffectiveness = (agility+ speed +strength+ willpower)) # creating a new object that includes the calculated variable combateffectiveness
head(avengers_effectiveness) #checking first rows to ensure variable was calculated and appears in data. 


```

3.  Creating dataset that only includes dead avengers with no superpowers in both spss and csv format.

```{r}
dead_avengers <- filter (avengers_effectiveness,superpower =="no" & died =="yes") #filter for avengers who have dead = yes and superpower =no 
head(dead_avengers) #check data to confirm filter worked

#write csv of new dataset
write.csv(dead_avengers, "dead_avengers.csv")

#write Sav (spss) using write_sav, part of haven package
write_sav(dead_avengers,"dead_avengers.sav")
```
